---
title: "S631 HW8"
author: "Shibi He"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
```

### ALR 8.2.1
```{r}
m1 = lm(Distance ~ Speed, data=stopping)
p1=inverseResponsePlot(m1, c(0, 0.5, 1, -1))
p1

boxCox(m1)
```


In the inverse fitted value plot, the best-fitting curve with $\hat{\lambda}=0.48$ and the square root curve for $\lambda=0.5$ is nearly identical, and so a square root transformation of Distance is suggested. Similarly, using the Box-Cox method, the square root transformation is in the 95% confidene interval for $\hat{\lambda}=0.48$, agreeing with the inverse fitted value plot. So the approporiate transformation for Distance is a square root transformation. 




### ALR 8.2.2
```{r}
with(stopping, invTranPlot(Speed, Distance))
```


The above graph shows that none of the fitted lines for $\lambda \in \{-1, 0, 1\}$ is close to the best-fitting line (i.e. $\hat{\lambda}=1.87$) and the RSS(1.87) is much lower than the RSS for the other three values. Therefore, none of these transformation is adequate. 




### ALR 8.2.3
```{r}
with(stopping, invTranPlot(Speed, Distance, 2))
```
 

Using $\lambda=2$ is almost identical to the best-fitting line ($\hat{\lambda}=1.87$) and RSS(2) is very close ot RSS(1.87), suggesting using a quadratic polynomial for the regressors. 
```{r}
m2 = lm(Distance ~ Speed + I(Speed^2), data=stopping)
summary(m2)
```

### ALR 8.2.4 

Hald (1960)'s model:
```{r}
m3 = lm(Distance ~ Speed + I(Speed^2), data=stopping, weights= 1/Speed^2 )
summary(m3)$coefficients
```


Square root transformation of Distance: (model in 8.2.1)
```{r}
m4 = lm(sqrt(Distance) ~ Speed, data=stopping)
summary(m4)$coefficients
```


Plot the fitted values: 
```{r}
# Plot fitted curve from Hald's model
newdata = data.frame(Speed=stopping$Speed, fitted.dist=fitted(m3))
newdata = newdata[order(newdata$Speed), ]

# Plot fitted value from square root transformation of distance
newdata2 = data.frame(Speed=stopping$Speed, fitted.dist=fitted(m4)^2)
newdata2 = newdata2[order(newdata2$Speed), ]

plot(Distance ~ Speed, stopping)
lines(newdata, col="red", lwd=2)
lines(newdata2, col="blue", lwd=2)
```



The fit of these two models are almost identical as these two models are essentially the same. They both fit the data very well. 


### ALR 8.6.1
```{r}
Wool$len=factor(Wool$len)
Wool$amp=factor(Wool$amp)
Wool$load=factor(Wool$load)
scatterplotMatrix(Wool, smooth = F,
                   diagonal = F)

#pairs(~cycles+len+amp+load, Wool)
```


These scatter plots show that as $\it{len}$ increases (i.e.moving from level 1 to 3), $\it{cycles}$ increases. As $\it{amp}$ and $\it{load}$ increases (i.e. moving from level 1 to 3), $\it{cycles}$ decreases. Moreover, the variation in $\it{cycles}$ also increases with $\it{len}$ and decreases with $\it{amp}$ and $\it{load}$.


### ALR 8.6.2
```{r}
m1 = lm(cycles ~ len + amp + load + len:amp + len:load + amp:load, data=Wool)
summary(m1)$coefficients

# Alternatively,
# m1=lm(cycles ~ (len+amp+load)^2, Wool)

```


```{r}
plot(Effect(c("len", "amp"), m1), rug=FALSE, grid=TRUE, multiline=TRUE)
```


The effects plot suggests that as $\it{len}$ moves from level 1 to level 3 (i.e. increases from 250mm to 350mm), the expected $\it{cycles}$ for all three different $\it{amp}$ increases. Moreover, for each specific $\it{len}$, as $\it{amp}$ moves from level 1 to level 3 (i.e. increases from 8mm to 10mm), the expected $\it{cycles}$ decreases. 


```{r}
# anova(m1)
# Anova(m1)
# Anova(m1, type=3)
```


### ALR 8.6.3

Fit the first-order mean function:
```{r}
m2 = lm(cycles ~ len + amp + load, data=Wool)
summary(m2)$coefficients
```


Use Box-Cox method to select a transformation for $\it{cycles}$:
```{r}
boxCox(m2)

# alternatively
# summary(powerTransform(m2))
```


The above graph shows that $\lambda=0$ is within the 95% confidence interval of the $\hat{\lambda}$, suggesting that a log transformation of the response $\it{cycles}$ might be adequate. 




### ALR 8.6.4

In the transformed scale, fit both the first-order and second-order model:
```{r}
# first-order mean model
model1 = lm(log(cycles) ~ len + amp + load, data=Wool)
summary(model1)$coefficients
```


```{r}
# second-order model
model2 = lm(log(cycles) ~ len + amp + load + len:amp + len:load + amp:load, data=Wool)
summary(model2)$coefficients
```


Compute an F-test comparing these two models:
```{r}
anova(model1, model2)
```


The F-statistic=2.216 and the pvalue=0.1325, suggesting that we cannot reject the null hypothesis that model1 (i.e. the first-order model) is adequate. 


```{r}
plot(Effect(c("len", "amp"), model1), rug=FALSE, grid=TRUE, multiline=TRUE)
```


Redraw this effects plot with cycles rather than log(cycles):
```{r}
# model3 = lm(cycles ~ len + amp + load, data=Wool)
# summary(model3)$coefficients
# 
# plot(Effect(c("len", "amp"), model3), rug=FALSE, grid=TRUE, multiline=TRUE)

plot(Effect(c("len", "amp"), model1, transformation=list(inverse=exp)), multiline=TRUE, rescale.axis=FALSE, ylab="cycles")
```



In the effects plot of Problem 8.6.2, the lines are not parallel because when we include the interaction term between $\it{len}$ and $\it{amp}$ into the model, when $\it{amp}$ moves from one level to another, its effects on expected $\it{cycles}$ is different for different velue of $\it{len}$. 


However, in this model without interaction terms, the lines in the effects plot are completely parallel. This is because the effects of $\it{amp}$ moving from one level to another are constant, reglardess of the value of $\it{len}$. 





